---
title: "Homework 4"
author: Evan Walker
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
setwd('D:\\Mathemagic\\Fall2019\\CS614_Interactive_Data_Analysis\\Classwork\\Classwork_7_lecture_9')
rm(list=ls())
library('DAAG')
require(ggplot2)
library('lme4')
library(MEMSS)
df1=ant111b
df2=Gun
df3=MathAchieve
load('pfadata.RData')
df4=pfa_all
```

## Problem 1
On Slide 13 of the Lecture 9 slides, we performed a random intercept model on the ant111b data,
with the only fixed effect being the intercept, i.e., average. Repeat this analysis eight times, where
within each reproduction, you eliminate one of the sites. For each of these runs, calculate the variance
partitioning coefficient, i.e. 𝜎𝜎𝐿𝐿Can you explain the results that you observe? The figure on Slide
6 may help your interpretation. What do you suppose would happen if instead of removing one site
at random, you simply removed one plot at random from each of the 8 sites?
(EACH TIME WE RUN WITH 7)
```{r, echo=FALSE}
summary(df1)
#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'DBAN'),]))
vpc1=1.6437/(1.6437+0.8049)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'LFAN'),]))
vpc2=1.6709/(1.6709+0.7415)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'NSAN'),]))
vpc3=1.3647/(1.3647+0.7619)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'ORAN'),]))
vpc4=1.2086/(1.2806+0.7985)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'OVAN'),]))
vpc5=1.6653/(1.6653+0.6384)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'TEAN'),]))
vpc6=1.5720/(1.5720+0.7988)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'WEAN'),]))
vpc7=1.5814/(1.5814+0.7479)

#summary(lmer(harvwt ~ 1+(1|site),data=df1[which(df1$site != 'WLAN'),]))
vpc8=1.5428/(1.5428+0.7744)

vpc=c(vpc1,vpc2,vpc3,vpc4,vpc5,vpc6,vpc7,vpc8)
names(vpc)=c('-DBAN','-LFAN','-NSAN','-ORAN','-OVAN','-TEAN','-WEAN','-WLAN')
vpc
print(paste("mean VPC:",mean(vpc)))
mean(vpc)
```
on average, 66% of the variance is due to between site variance. When removing site "OVAN", we get the highest intra-site variance of 72 percent,which would imply that "OVAN" varies little from other sites. Likewise, the lowest VPC is "ORAN" with 58% intra-site variance implying that "ORAN" differed significantly from other sites.

If we instead removed one plot at random, I do not believe our results would differ significantly since we would expect the yield of a plot at a site to be similar to other plots on the same site, but not similar to plots on other sites. Hence preserving the variation between sites in our analysis.

## Problem 2
The data set Guns from the MEMSS package reports the number of rounds fired per minute, by each
of nine teams of gunners, each tested twice, once each with two different methods. Of the nine teams,
three were comprised of men with slight builds, three with average builds, and three with heavy
builds. Does the number of rounds fired significantly vary based on build or firing method? What
proportion of the total variance is due to between-team versus within-team considerations, i.e. what
is the variance portioning coefficient?
```{r, echo=FALSE}
summary(df2)
#summary(aov(rounds~Error(Physique),data=df2))
#summary(aov(rounds~Error(Method),data=df2))
#summary(aov(rounds~Physique,data=df2)) #not sig
#summary(aov(rounds~Method,data=df2))  #sig

summary(aov(rounds~Error(Team),data=df2))
summary(lmer(rounds ~ 1+(1|Team),data=df2))
sigW=4.662
sigB=0
vpc=sigB/(sigB+sigW)   #none of the variation is due to differences between team
```
based on the ANOVA, it seems that rounds fired significantly differ with respect to Method, but not Physique.
As for the between team variance, the VPC indicates that there is no significant differences between teams.


## Problem 3
The data set MathAchieve describes math achievement scores from various schools. The binary
variables Minority and Sex and the continuous variable SES are obviously fixed-effects variables.
However, the School variable can be chosen as either fixed or random effect. Discuss how different
purposes for the study might result in you treating School as either a fixed or random effect. Carry
out the random effect analysis and detail the results, in terms of the significance of the fixed effects
and the contributions of variances. Which model better fits the data, a random-intercept or random
slope model?
```{r, echo=FALSE}
summary(df3)
m1=lmer(MathAch ~ Sex+Minority+SES+(1|School),data=df3)
#summary(lm(MathAch ~ Sex+Minority+SES+School,data=df3))
m2=lmer(MathAch ~ Sex+Minority+SES+(MEANSES|School),data=df3)
anova(m1,m2)
```
It makes the most sense to describe math achiecements with a random intercept WTR to School. I'm not entirely sure what variable to make the random effect of slope. A random slope model may be benificial for if we believe that the change in math achievements vary with respect to school.


## Problem 4
Repeat the multiple baseline model that was outlined on Slide 49, but allow for change in slope as a
result of the intervention in addition to change level. Do random slopes make sense? Summarize your
results.
```{r, echo=FALSE}
#random intercept, fixed slope (time is not a preditor variable)
m3=lmer(meancounts ~ IVflag+(1|HomeID),data=df4,REML=F)
summary(m3)
#ranef(m3)

#random intercept+slope, fixed slope (time is not a preditor variable)
m4=lmer(meancounts ~ IVflag+(IVflag|HomeID),data=df4,REML=F)
summary(m4)
#ranef(m4)
anova(m3,m4)
```
Based on AIC and BIC, the random slope and random intercept model is preferred. random slopes make sense since there is not 1 single baseline, but each home has its own baseline hence should be interpretted as acting differently per home


